---
title: "Moving along with R"
description: |
  As I've progressed further with learning R and R Studio, I now am quite confident in my abilities. 
author:
  - name: Jayde Lyon
    url: https://www.google.com
date: 04-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#first load the libraries we'll need
library(tidyverse)
library(plotly)
library(nycflights13) # <- make sure to install this package 


```

## Functions

1. Create a function that will return the sum of 2 numbers.

```{r}

sum_of_2 <- function(number1, number2){
  
  result <- number1+number2
    
    return(result)
}


```


2. Create a function that will return the mean of a vector of numbers fed to it.  
A sample vector of values is provide here to help.

```{r}

mynumbers <- sample(1:100, 10)

mean_of_vec <- function(numbers){
  
  result <- mean(numbers)
    
    return(result)
}

mean_of_vec(mynumbers)


```


3. Now, expand on the function you made above, and instead of just returning the number, have it return the sentence:  
"The mean is ___" where the blank is the calculated result.

```{r}

mean_of_vec <- function(numbers){
  
  result <- paste0("The mean is ", mean(numbers))
    
    return(result)
}

mean_of_vec(mynumbers)



```


```{r}
#load the data to use for the rest of the assignment questions
flights <- nycflights13::flights

```

4. Using the `flights` dataset from the package `nycflights13` loaded in the above cell, create a function that filters the data frame to only include records from a single originating airport (column name `origin`).

```{r}

origin_flight <- function(airport){
  
  result <- flights %>% 
            filter(origin == airport)
    
    return(result)
}

origin_flight('JFK')

```


5. Similarly, using the `flights` dataset, create a function that:  
- filters the data frame to only include records from a single originating airport (column name `origin`), and then  
- aggregates the results by airline (`carrier`) to show how many flights each airline has from that airport.

```{r}


origin_airline <- function(airport){
  
  result <- flights %>% 
            filter(origin == airport) %>% 
            count(carrier)
    
    return(result)
}

origin_airline('JFK')

```


6. Similarly, using the `flights` dataset, create a function that:  
- filters the data frame to only include records from a single originating airport (column name `origin`), and then  
- filter to only include flights longer than a specific number of minutes (`air_time`), and then  
- aggregates the results by airline (`carrier`) to show how many flights of that length each airline has from that airport.

```{r}

origin_airtime <- function(airport, minutes){
  
  result <- flights %>% 
            filter(origin == airport) %>% 
            filter(air_time >= minutes) %>% 
            count(air_time)
    
    return(result)
}

origin_airtime('JFK', 100)


```



## Plotly charts

7. Using the `flights` dataset, use plotly to create a bar chart showing the breakdown of flights by airline that originate
from JFK airport (`origin`) and are more than 400 minutes in length (`air_time`).

```{r}

#data prep

JFK_origin <- flights %>% 
  filter(origin == 'JFK') %>% 
  filter(air_time >= 400) %>% 
  count(carrier)

```

```{r}

JFK_flights <- plot_ly(data= JFK_origin, x = ~carrier, y = ~n, type = 'bar') %>% 
  layout(title = "Flights From JFK Over 400 Minutes",
         xaxis = list(title = "Carrier"),
         yaxis = list (title = "Number of Flights"))

JFK_flights

```


8. Using the `flights` dataset, use plotly to create a bar chart showing which destination airports (`dest`) have the most
flights with durations of over 400 minutes (`air_time`).

```{r}

# data prep

dest_400 <- flights %>% 
  filter(air_time >= 400) %>% 
  count (dest)


```

```{r}

dest_flights <- plot_ly(data= dest_400, x = ~dest, y = ~n, type = 'bar') %>% 
  layout(title = "Destinations of Flights With Over 400 Minute Flights",
         xaxis = list(title = "Destination"),
         yaxis = list (title = "Number of Flights"))

dest_flights
```

9. Using the `flights` dataset, use plotly to create a donut chart showing the share of flights each airline has for 
flights with a destination of Daniel K. Inouye International Airport (HNL) in Honolulu.

```{r}

# data prep

HNL_carrier <- flights %>% 
              filter(dest == 'HNL') %>% 
            count(carrier)


```

```{r}

HNL_donut <- HNL_carrier %>% plot_ly(labels = ~carrier, values = ~n) %>% 
                add_pie(hole = 0.6) %>% 
                layout(
                  title = "Share of Airline Flights to HNL",  showlegend = F,
                  xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                  yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


HNL_donut
```

10. Using the `flights` dataset, use plotly to create a scatterplot chart for flights with a destination of 
Daniel K. Inouye International Airport (HNL) in Honolulu, that shows the relationship between the departure delay (`dep_delay`) and arrival delay (`arr_delay`).

```{r}

# data prep

HNLdest <- flights %>% 
          filter(dest == 'HNL') 


```

```{r}

plot_ly(data = HNLdest, x = ~dep_delay, y = ~arr_delay,
        marker = list(size = 10,
                             color = 'blue',
                             line = list(color = 'blue',
                                         width = .5)))

