---
title: "My first R assignment"
description: |
 This is my first R assignment, looking at the election results and which members of Congress voted to impeach former President Donald J. Trump. 
author:
  - name: Jayde Lyon
    url: https://www.google.com
date: 04-19-2021
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - Class work
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#first we'll load our libraries
library(tidyverse)
library(lubridate)

# run this to load the data for this assignment
# it will create a dataframe called "impeach," with all House Democrats and election results/demographics
# it indicates whether or not the member publicly called for the impeachment 
# note: THIS IS FOR THE FIRST TRUMP IMPEACHMENT IN 2020, NOT THE SECOND IN 2021
impeach <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRh8d5JaDqBtByzNw2JZF3idaACobDhhk-p7chJoktA0HawELHFOvjQOqCpzGA4MGONvPlR7GASqW-K/pub?gid=1765341510&single=true&output=csv")


```


FOR EACH OF THE QUESTIONS BELOW IN EACH CODE CHUNK, WRITE YOUR WORKING R CODE 
TO RETURN THE REQUESTED RESULTS

```{r}
# 1) The column "for_impeachment" indicates whether the member has publicly called for
# an impeachment inquiry. Filter to return only the ones where the answer is NO.    

impeach %>% 
  filter(for_impeachment == 'NO')



```


```{r}
# 2) Filter to return only results where a member is both against impeachment, and comes from a 
# district that President Trump won in 2016 (which is noted in the "p16winningparty" column)


impeach %>% 
  filter (for_impeachment == 'NO', p16winningparty == 'R')


```


```{r}
# 3) Filter for only results where a member is against impeachment, comes from a 
# district that President Trump won in 2016 (which is noted in the "p16winningparty" column),
# and also comes from a district that Mitt Romney won in 2012 ("p12winningparty").

impeach %>% 
  filter (for_impeachment == 'NO', p16winningparty == 'R', p12winningparty == 'R')


```


```{r}
# 4) Filter for only results from September 2019 where a member is a YES for impeachment. 

impeach %>% 
  filter(date_month == '9', date_year == '2019', for_impeachment == 'YES')



```


```{r}
# 4) Filter for only results where a member is a YES for impeachment and is from a district
# where Clinton won more than 70 percent of the vote in 2016 (found in column "clinton_percent")

impeach %>% 
  filter(clinton_percent >= '70.0', for_impeachment == 'YES')


```


```{r}
# 5) Sort the entire dataframe based on the percentage of a district that has a 
# bachelor's degree or higher ("pct_bachelors"), from lowest to highest


impeach %>% 
  arrange (pct_bachelors)


```


```{r}
# 6) Sort the just those who are NO on impeachment based on the percentage of a district that has a 
# bachelor's degree or higher ("pct_bachelors"), from lowest to highest

impeach %>% 
  filter(for_impeachment == 'NO') %>% 
  arrange(pct_bachelors)



```


```{r}
# 7) Sort the just those who are NO on impeachment based on the percentage of a district that has a 
# bachelor's degree or higher ("pct_bachelors"), from lowest to highest.
# Then filter those records by only those whose bachelor's percentage is below the national average (found
# in the "pct_bachelors_compared_to_national" column).

no_impeach <- impeach %>% 
  filter(for_impeachment == 'NO') %>% 
  arrange(pct_bachelors)

no_impeach %>% 
  filter(pct_bachelors_compared_to_national == 'BELOW')



```


```{r}
# 8) Filter for only members from New Jersey who are NO on impeachment


impeach %>% 
  filter(for_impeachment == 'NO', state == 'NJ')


```


```{r}
# 9) Filter for those who were YES on impeachment, with a declared date prior to 2019. So only
# those with dates before 2019.  Then sort those so that the highest Clinton vote percentages are 
# at the top.   

impeach %>% 
  filter (date_year < '2019', for_impeachment == 'YES') %>% 
  arrange (desc(clinton_percent))


```


```{r}
# 10) Answer this question using grouping.
# How many members in the dataset who are holdouts on impeachment
# come from districts with a GDP that is above the national figure? How many below?

impeach %>% 
  filter (for_impeachment == 'NO') %>% 
  count(gdp_above_national)


```


```{r}
# 11) Answer this question using grouping.
# Similar to #10 above, now break down both those who favor impeachment AND who don't
# and show which come from district above or below national GDP


impeach %>% 
  count (for_impeachment, gdp_above_national)



```



```{r}
# 12) Answer this question using grouping.
# Now do the same as #11, but this time instead of GDP look at above/below 
# the national avg percent of college graduates (pct of pop with bachelor's degrees)


impeach %>% 
  count (for_impeachment, pct_bachelors_compared_to_national)


```


```{r}
# 13) Answer this question using grouping.
# Instead of showing the above/below national average counts for those for or 
# against impeachment, this time show the combined MEAN/AVERAGE percentage of college
# grads for all districts that are Yes for impeachment vs. No

impeach %>% 
  group_by(for_impeachment) %>% 
  summarize(mean(pct_bachelors)) 



```



```{r}
# 14) Answer this question using grouping.
# Do the same as #13, but this time show the combined MEAN/AVERAGE percentage of 
# the vote that Donald Trump received for all districts that are Yes for 
# impeachment vs. No


impeach %>% 
  group_by(for_impeachment) %>% 
  summarize(mean(trump_percent)) 

